---
title: "Creating a Bespoke Model"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Creating a Bespoke Model}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(bespoke)
library(modeldata)
```

Sometimes a simple series of `if` statements are enough for a subject matter expert to make a fairly good prediction from input data.
In such cases, it can be helpful to hand-craft a function that can make simple predictions on new data.
However, it is not easy to put such functions on the same footing as actual models for comparison.

With `{bespoke}`, you can construct "fitted models" which behave the same as any other model within the `{tidymodels}` framework, but which use hand-crafted functions for prediction. Here we will demonstrate a simple case.

## Oil Data

We will work with the `oils` data from `{modeldata}`.
This dataset consists of 96 samples of commercial oils, each of which belongs to one of 7 classes of oil: corn, olive, peanut, pumpkin, rapeseed, soybean, or sunflower.

```{r oil-data}
data(oils)
summary(oils)
```

We'll construct a function to choose a class from some simple properties in the input data.

## Our Hand-Crafted Model Function

For now, we will simply return a random choice for each row of data.
As we develop this package further we'll create a more realistic use case.

```{r hand-crafted}
random_baseline <- function(new_data, n_classes) {
  # Return a number between 1 and n_classes for each row of input.
  return(
    sample(seq_len(n_classes), nrow(new_data))
  )
}
```

## "Training" a Model

In this case we need to specifically tell the "model" how many classes we have.
We will likely make this a standard parameter available to any bespoke functions in the future.

```{r oil-fit}
oil_fit <- bespoke_class(
  class ~ ., 
  oils, 
  fn = random_baseline, 
  n_classes = 7
)
```

That object will now behave like other model objects!

## Using Our Model

```{r oil-predict}
oils_no_classes <- oils[, 1:7]
head(oils_no_classes)
predict(oil_fit, new_data = head(oils_no_classes))
```

Note that in this case the predictions are completely random.

```{r oil-predict-again}
predict(oil_fit, new_data = head(oils_no_classes))
```
